A Bayesian Perspective \\\large{Extracorporeal CPR for Refractory Out-of-Hospital Cardiac Arrest}

### 2023-02-12 manuscript/manuscript.qmd
getting closer

The INCEPTION researchers have ddressed an important clinical question in the most challenging of research environments and are to be congratulated on their trial design, its execution, and a nuanced discussion. However the natural constraints of standard statistical analyses, also known as null hypothesis significance testing, limits the quantitative appreciation of their data and prevents a full and comprehensive exploitation and updating of past knowledge. For trials such as INCEPTION that fail to meet statistical significance and are often thought of as "negative" trials, null hypothesis significance testing favors confusion "between absence of evidence and evidence of absence"[@RN6765]. Rather than dichotomizing results into statistical significance or not, with its obligatory loss of information, Bayesian analysis concentrates on estimation of key outcome differences with direct probability measures of their uncertainty.

Although this  Bayesian analysis presents the strengths noted above, it also had limitations. First, we did not have access to either
individual data from any of the trials limiting the possibility of examining specific subgroups. Also for the per protocol analysis, this data was only available for survival and not the primary outcome that included survival with intact neurological status. The threshold choices for superiority and equivalence may be seen as arbitrary, although the analyses can be easily repeated for different choices. Given the existence of only 3 trials  estimating the very divergent baseline cCPR rates could not be reliably modeled. We choose not to perform a traditional Bayesian meta-analysis with
a semi-informative prior for the between-study variation  because our
goal was to illustrate how Bayesian principles could be informative
when applied to the analysis and interpretation of new trial data, both in the situation of no and considerable previous knowledge.

The overall message from this study is that standard statistical analyses leading to a conclusion of "similar survival effects" may be overly simplified and potentially inaccurate. This Bayesian analysis demonstrates clearly that while any definitive conclusions regarding the superiority of eCPR are lacking, the possibility of a clinically meaningful benefit, or even less likely the possibility of clinically meaningful harm, has not been reasonably excluded and continued research is necessary to clarifying the remaining uncertainties. The final dilemma confronting clinicians is that even a Bayesian analysis of a randomized trial provides only probability estimates of an average treatment effect and not the elusive individual treatment effect.

>>>>

The INCEPTION researchers have addressed an important clinical question in the most challenging of research environments and are to be congratulated on their trial design, its execution, and a nuanced discussion. However the natural constraints of standard statistical analyses, also known as null hypothesis significance testing, limits the quantitative appreciation of their data, and prevents a full and comprehensive exploitation and updating of past knowledge. For trials such as INCEPTION that fail to meet statistical significance and are often incorrectly thought of as "negative" trials, null hypothesis significance testing favors confusion "between absence of evidence and evidence of absence"[@RN6765]. Rather than dichotomizing results into statistical significance or not, with an obligatory loss of information, Bayesian analysis concentrates on estimation of key outcome differences with direct probability measures of their uncertainty.

Although this  Bayesian analysis presents the strengths noted above, it also had limitations. First, we did not have access to 
individual data from any of the trials limiting the possibility of examining specific subgroups. Also for the per protocol analysis, this data was only available for survival and not the primary INCEPTION outcome that included survival with intact neurological status. The threshold choices for superiority and equivalence may be seen as arbitrary, although the analyses can be easily repeated for different choices. Given the existence of only 3 trials  estimating the very divergent baseline cCPR rates could not be reliably modeled. We choose not to perform a traditional Bayesian meta-analysis with
a semi-informative prior for the between-study variation  because our
goal was to illustrate how Bayesian principles could be informative
when applied to the analysis and interpretation of new trial data, both in the situations of no and considerable previous knowledge.

The "take home" message from this study is that standard statistical analyses resulting in a conclusion of "similar survival effects" may be overly simplified and potentially inaccurate. This Bayesian analysis demonstrates that while any definitive conclusions regarding the superiority, inferiority or equivalence of eCPR to cCPR are lacking. Rather the possibility of a clinically meaningful benefit, or even less likely the possibility of clinically meaningful harm, has not been reasonably excluded and continued research is necessary to clarifying the remaining uncertainties. The final dilemma confronting clinicians is that even Bayesian analyses of randomized trials provide only probability estimates for an average treatment effect and not for the elusive individual treatment effect.

### 2023-02-12 manuscript/manuscript.tex
getting closer

\begin{keyword}
    extracorporeal CPR \sep 
    Bayesian statistics
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, interior hidden, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

The INCEPTION researchers have ddressed an important clinical question
in the most challenging of research environments and are to be
congratulated on their trial design, its execution, and a nuanced
discussion. However the natural constraints of standard statistical
analyses, also known as null hypothesis significance testing, limits the
quantitative appreciation of their data and prevents a full and
comprehensive exploitation and updating of past knowledge. For trials
such as INCEPTION that fail to meet statistical significance and are
often thought of as ``negative'' trials, null hypothesis significance
testing favors confusion ``between absence of evidence and evidence of
absence''\citep{RN6765}. Rather than dichotomizing results into
statistical significance or not, with its obligatory loss of
information, Bayesian analysis concentrates on estimation of key outcome
differences with direct probability measures of their uncertainty.

Although this Bayesian analysis presents the strengths noted above, it
also had limitations. First, we did not have access to either individual
data from any of the trials limiting the possibility of examining
specific subgroups. Also for the per protocol analysis, this data was
only available for survival and not the primary outcome that included
survival with intact neurological status. The threshold choices for
superiority and equivalence may be seen as arbitrary, although the
analyses can be easily repeated for different choices. Given the
existence of only 3 trials estimating the very divergent baseline cCPR
rates could not be reliably modeled. We choose not to perform a
traditional Bayesian meta-analysis with a semi-informative prior for the
between-study variation because our goal was to illustrate how Bayesian
principles could be informative when applied to the analysis and
interpretation of new trial data, both in the situation of no and
considerable previous knowledge.

The overall message from this study is that standard statistical
analyses leading to a conclusion of ``similar survival effects'' may be
overly simplified and potentially inaccurate. This Bayesian analysis
demonstrates clearly that while any definitive conclusions regarding the
superiority of eCPR are lacking, the possibility of a clinically
meaningful benefit, or even less likely the possibility of clinically
meaningful harm, has not been reasonably excluded and continued research
is necessary to clarifying the remaining uncertainties. The final
dilemma confronting clinicians is that even a Bayesian analysis of a
randomized trial provides only probability estimates of an average
treatment effect and not the elusive individual treatment effect.

>>>>

\begin{keyword}
    extracorporeal CPR \sep 
    Bayesian statistics
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, interior hidden, enhanced, breakable]}{\end{tcolorbox}}\fi

@@

The INCEPTION researchers have addressed an important clinical question
in the most challenging of research environments and are to be
congratulated on their trial design, its execution, and a nuanced
discussion. However the natural constraints of standard statistical
analyses, also known as null hypothesis significance testing, limits the
quantitative appreciation of their data, and prevents a full and
comprehensive exploitation and updating of past knowledge. For trials
such as INCEPTION that fail to meet statistical significance and are
often incorrectly thought of as ``negative'' trials, null hypothesis
significance testing favors confusion ``between absence of evidence and
evidence of absence''\citep{RN6765}. Rather than dichotomizing results
into statistical significance or not, with an obligatory loss of
information, Bayesian analysis concentrates on estimation of key outcome
differences with direct probability measures of their uncertainty.

Although this Bayesian analysis presents the strengths noted above, it
also had limitations. First, we did not have access to individual data
from any of the trials limiting the possibility of examining specific
subgroups. Also for the per protocol analysis, this data was only
available for survival and not the primary INCEPTION outcome that
included survival with intact neurological status. The threshold choices
for superiority and equivalence may be seen as arbitrary, although the
analyses can be easily repeated for different choices. Given the
existence of only 3 trials estimating the very divergent baseline cCPR
rates could not be reliably modeled. We choose not to perform a
traditional Bayesian meta-analysis with a semi-informative prior for the
between-study variation because our goal was to illustrate how Bayesian
principles could be informative when applied to the analysis and
interpretation of new trial data, both in the situations of no and
considerable previous knowledge.

The ``take home'' message from this study is that standard statistical
analyses resulting in a conclusion of ``similar survival effects'' may
be overly simplified and potentially inaccurate. This Bayesian analysis
demonstrates that while any definitive conclusions regarding the
superiority, inferiority or equivalence of eCPR to cCPR are lacking.
Rather the possibility of a clinically meaningful benefit, or even less
likely the possibility of clinically meaningful harm, has not been
reasonably excluded and continued research is necessary to clarifying
the remaining uncertainties. The final dilemma confronting clinicians is
that even Bayesian analyses of randomized trials provide only
probability estimates for an average treatment effect and not for the
elusive individual treatment effect.
